Automated Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{NSF Grant ???} }

### 2021-03-31 paper/paper_new.tex
Updates from Overleaf
Finished a pass through Section 5.

\begin{abstract}
     Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections.
The metadata that is vital for subsequent machine learning, analysis and
scientific discovery based on these specimen image repositories
is often unavailable, sparse or incorrect.
As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.
Our goal is to develop image metadata generation methods that both support
the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.
\end{abstract}

@@

This limitation is magnified when trying get a handle on the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community advocated for data fitness standards \cite{chapman2020developing}. Even so, metadata properties used to record image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, justifying the capture of 22 metadata properties supporting selection of high quality images for the development of a bgnn. Although, the manual generation of this metadata is costly in terms of labor and time.
%will add cites above, depending on paper length.

@@

%\cite{heidorn2008automatic} speciment labels \cite{manso2004automatic} carteogrpahic info.

@@

Digitized specimen collections offer numerous outstanding opportunities
for new scientific studies and discoveries.

The amount of data is immense, with image collections containing 10,000's
of entries. These discoveries will only be possible
with computational techniques that can process this vast amount of data.

@@

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
\todo{Not sure if I can phrase it that way or not since I haven't looked at all the images to see if the ruler itself was found or something else was labeled a ruler} In 56 of the images, the ruler itself was found but the ``2'' and/or the ``3'' on it was not, so a scale calculation could not be performed, producing a 99.2\% success rate for being able to compute 
\verb|scale|.

\subsection{Side Detection}
For 246 of the images \verb|detectron| was unable to find a fish eye. Of the remaining \(6,998\) images where it did, the correct side (\verb|left| or \verb|right|) was detected in all but 6 cases.

\subsection{Clock Value}
Clock position values were successfully generated for \(6,991\) of the images. Of those, all but 8 were within \(\pm{}1\) of the correct result.

@@

\subsection{Brightness/Uniformity Results?}
\begin{verbatim}
Mean intensity in bounding box | Count

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/011196_pred}
  \includegraphics[width=0.49\linewidth]{images/011196_mask}
  \includegraphics[width=0.49\linewidth]{images/52399_pred}
  \includegraphics[width=0.49\linewidth]{images/52399_mask}
  \includegraphics[width=0.49\linewidth]{images/56885_pred}
  \includegraphics[width=0.49\linewidth]{images/56885_mask}
  \caption{Examples of masks and bounding boxes from detectron (left) and pixel analysis (right).}
\end{figure}
Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box. Of the 50 images reviewed manually for these properties, all mask and bounding boxes were correctly placed on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins. Specifically, 22 mask and bounding boxes covered the entire fish, 16 missed some of the tail, and 12 missed most or all of the tail.

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images. 50 of these images were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ.~\cite{imagejCite} The average error for the scale calculation was \(0.911\%\), and the average error for the fish length calculation was \(5.87\%\).

\section{Discussion}
\textbf{Todo}: Will need some sort of intro/framing paragraph
\subsection{Results}\todo{Just going to put the results discussion in small sections and we can change it later if that doesn't make sense}
\subsubsection{Object Detection}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/none1}
  \includegraphics[width=0.49\linewidth]{images/double1}
  \caption{An example of a fish that was not detected (left) and a fish that was detected twice (right).}
\end{figure}

@@

\section*{Acknowledgment}

Vishal? Hank's undergrads? Other team members?

>>>>

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
     Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been acquired worldwide by museums and other research institutions, undertakings that have produced large image collections.
The metadata that is vital for subsequent machine learning, analysis and
scientific discovery based on these specimen image repositories
is often unavailable, sparse or incorrect.
As a step towards improving metadata in specimen research image collections, our team is developing methods for automatically analyzing fish images to extract a variety of important features. These fish specimens are being studied for a larger project entitled Biology Guided Neural Networks (BGNN), which is developing a novel class of artificial neural networks that can exploit the machine readable and predictive knowledge about biology that is available in specimen images, phylogenies and anatomy ontologies. Using a combination of machine learning and image informatics tools and techniques, we can accurately determine metadata such as fish quantity and location within images, fish orientation and other quantitative fish features, image scaling based on ruler identification and measurement, and general image quality metrics for a substantial number of the images being used in the BGNN project.
Our goal is to develop image metadata generation methods that both support
the research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing quantitative features in support of specific biological investigations.
\end{abstract}

@@

This limitation is magnified when trying get a handle on the actual quality of the digitized specimen. Descriptive oriented standards support search and retrieval, and the biodiversity community advocated for data fitness standards \cite{chapman2020developing}. Even so, metadata properties used to record image quality is severely limited and generally missing. This point is addressed in detail in \cite{leipzig2021biodiversity}, justifying the capture of 22 metadata properties supporting selection of high quality images for the development of a bgnn. Although, the manual generation of this metadata is costly in terms of labor and time.

@@

%\cite{heidorn2008automatic} speciment labels \cite{manso2004automatic} cartographic info.
%above paragraph..need to finesse a couple of sentences w/results, weave in these 2 citations, and then say we can look to image analysis.

@@

Digitized specimen collections offer numerous outstanding opportunities
for new scientific studies and discoveries.
%I think we can go from the above, statement, directly to the paragraph that starts with "Given the many..." Then, specifically state the goal of this paper is to XXXX address the need, or the goal is to investigate, test, (pick your verb) an approach automatically generating metadata, combining off-the-shelf software and custom written, or is it simply test off-the-shelf software, and later the realization that customization happens.
The amount of data is immense, with image collections containing 10,000's
of entries. These discoveries will only be possible
with computational techniques that can process this vast amount of data.

@@

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
% \todo{Not sure if I can phrase it that way or not since I haven't looked at all the images to see if the ruler itself was found or something else was labeled a ruler}
In 56 of the images, the ruler itself was found but the ``2'' and/or the ``3'' on it was not, so a scale calculation could not be performed, producing a 99.2\% success rate for
the \verb|scale| computation.

\subsection{Side Detection}
\verb|detectron| was unable to find a fish eye
in 246 of the images. Of the remaining \(6,998\) images, the correct side
(\verb|left| or \verb|right|) was detected in all but 6 cases, producing
a 96.5\% correct rate.
\todo{Comment on the 6 cases. What failed?}

\subsection{Clock Value}
Clock position values were successfully generated for \(6,991\) of the images. Of those, all but 8 were within \(\pm{}1\) of the correct result,
our definition of a correct/acceptable result. The correct rate for this
computation is 96.4\%.
\todo{Comment on the cases where we were unable to compute clock value. What failed?}

@@

\subsection{Brightness/Uniformity Results?}
\textbf{Todo}: Need standard deviations here, too. To show that the
differences are significant and meaningful. Clearly the trend is there.\newline
It looks like we should leave out Uniformity.

\begin{verbatim}
Mean intensity in bounding box | Count

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.49\linewidth]{images/011196_pred}
  \includegraphics[width=0.49\linewidth]{images/011196_mask}
  \includegraphics[width=0.49\linewidth]{images/52399_pred}
  \includegraphics[width=0.49\linewidth]{images/52399_mask}
  \includegraphics[width=0.49\linewidth]{images/56885_pred}
  \includegraphics[width=0.49\linewidth]{images/56885_mask}
  \caption{Examples of masks and bounding boxes from detectron (left) and pixel analysis (right).}
\end{figure}

\subsection{Mask and Bounding Box}
Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box.
50 randomly-chosen images were reviewed manually to evaluate the
calculation.  All 50 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 22 masks and bounding boxes covered the entire fish,
16 missed some of the tail, and 12 missed most or all of the tail.

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
\todo{What happened the excluded images?}
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same(?) 50 test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ.~\cite{imagejCite}
In this subset of images the average error for the scale calculation
was \(0.91\%\), and the average error for the fish length calculation was \(5.87\%\).

\section{Discussion}
\textbf{Todo}: Will need some sort of intro/framing paragraph
%DRAFT from jg: Our results provide insight into how off-the-shelf software, supplemented by additional XXX [programming?] can support automatic metadata generation. Results are promising in a number of respects, helping us to identify more clearly areas, where more work is needed [e.g., failed fish, failed rulers]
%Discussion needed to weave in at least a couple of citations. Happy to discuss
\subsection{Results}\todo{Just going to put the results discussion in small sections and we can change it later if that doesn't make sense}
\subsubsection{Object Detection}
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/none1}
  \includegraphics[width=0.49\linewidth]{images/double1}
  \caption{An example of a fish that was not detected (left) and a fish that was detected twice (right).}
\end{figure}

@@

\section*{Acknowledgment}
%can make possible note about larger BGNN team, and/or general note to Tulane data in-putters for giving us data to compare, and/or INHS collection. [we can discuss]
